---
title: "set_10_Volatility1"
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---

Set data directory path
```{r setup, include=FALSE, cache = FALSE}
if (!require("knitr")) install.packages("knitr")
library(knitr)
opts_knit$set(root.dir = "C:/Teaching/Undergrad/data ECO374")
```
 
Install and load required packages
```{r}
chooseCRANmirror(graphics=FALSE, ind=33)
if (!require("MASS")) install.packages("MASS")
if (!require("moments")) install.packages("moments")
if (!require("MTS")) install.packages("MTS")
install.packages("MFTSR", repos="http://R-Forge.R-project.org")

library(aTSA)
library(xts)
library(anytime)
library(MASS)
library(moments)
library(MTS)
library(MFTSR)
```

Dowload data on S&P500 index, daily frequency
Source: https://finance.yahoo.com/quote/%5EGSPC/history?p=%5EGSPC

Load, extract Close price, convert to xts format, and plot
```{r}
setwd("C:/Teaching/Undergrad/data ECO374")
 
data_SP500_full <- read.csv(file="SP500_daily.csv", header=TRUE, sep=",")
 
xtsdata_SP500 <- xts(x=data_SP500_full[c("Close")], order.by=anytime(data_SP500_full$Date))
 
plot.xts(xtsdata_SP500, plot.type = c("single"), main ="S&P 500 index, close price" , legend.loc = "topleft", grid.col="lightgray")
```

Plot of returns
```{r}
xtsdata_SP500_returns <- diff(log(xtsdata_SP500))
xtsdata_SP500_returns <- na.omit(xtsdata_SP500_returns)

plot.xts(xtsdata_SP500_returns, plot.type = c("single"), main ="S&P 500 index, returns" , legend.loc = "topleft", grid.col="lightgray", col="blue")
```

Summary statistics of returns
```{r}
summary(xtsdata_SP500_returns)
```

Note that the mean of the returns is very close to zero, and hence returns SQUARED closely approximate
the VARIANCE of the returns.

Plot of returns SQUARED (variance)
```{r}
xtsdata_SP500_returns_SQUARED <- xtsdata_SP500_returns^2

plot.xts(xtsdata_SP500_returns_SQUARED, plot.type = c("single"), main ="S&P 500 index, returns SQUARED (variance)" , legend.loc = "topleft", grid.col="lightgray", col="red")
```

ACF plots
```{r}
maxlag <- 30
ACF1 <- acf(as.matrix(xtsdata_SP500_returns$Close), main=NULL, lag.max=maxlag, plot=FALSE)
ACF2 <- acf(as.matrix(xtsdata_SP500_returns_SQUARED$Close), main=NULL, lag.max=maxlag, plot=FALSE)

par(mfcol=c(1,2), mai=c(0.4, 0.4, 0.7, 0.1))
plot(ACF1[1:maxlag],ylim=c(-0.05,0.25), main="ACF returns")
plot(ACF2[1:maxlag],ylim=c(-0.05,0.25), main="ACF returns SQUARED (variance)")
```

Unit root test for returns
```{r}
adf.test(xtsdata_SP500_returns)
```


Plot histogram of returns with an overlay of best-fit Gaussian density 
Note that the returns do not fit the Gaussian density well
```{r}
fit <- fitdistr(xtsdata_SP500_returns, "normal")
para <- fit$estimate
hist(xtsdata_SP500_returns, prob = TRUE, breaks=200, xlim=c(-0.03,0.03), col="lightgrey")
curve(dnorm(x, para[1], para[2]), col = 2, add = TRUE)
```

Moments of the density of returns
```{r}
X <- xtsdata_SP500_returns$Close
all.moments( X, central=TRUE, order.max=4 )
```


The returns have a leptocurtic density with "heavy tails"
```{r}
hist(xtsdata_SP500_returns, prob = TRUE, breaks=1000, ylim = c(0, 10), xlim=c(-0.04,-0.02), col="lightgrey")
curve(dnorm(x, para[1], para[2]), col = 2, add = TRUE)
```



EWMA: estimate the smoothing parameter lambda
If lambda is negative, then the mul-tivariate Gaussian likelihood is used to estimate the smoothing parameter.
```{r}
Y <- as.matrix(xtsdata_SP500_returns$Close)
EWMA <- EWMAvol(Y, lambda = -1)
```

EWMA: plot
```{r}
EWMA <- ewmaVol(xtsdata_SP500_returns, lambda=0.941)
plot(fitted(EWMA))
```


